ms2

### 2017-08-30 ms2/ms2_display.Rmd
ms2 analyze, display

```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
ineq = read_tsv('ineq.tsv')

ineq_summary = ineq %>%
  group_by(year, unit_type, unit, drug_group) %>%
  summarize(mean=weighted.mean(n_claims, w=n_bene),
            fnz=sum(n_bene[n_claims > 0]) / sum(n_bene),
            mup=mean/fnz) %>%
  group_by(unit_type, unit, drug_group) %>%
  summarize_at(vars(mean, fnz,mup), mean) %>%
  ungroup()

@@

ineq %>%
  filter(unit_type=='state') %>%
  # sum across units
  group_by(year, drug_group, n_claims) %>%
  summarize(n_bene=sum(n_bene)) %>%
  ungroup() %>%
  # compute the number of claims made by people making 1 claim, etc.
  mutate(n_class_claims = n_claims * n_bene) %>%
  # compute percentages: fr. of benes w/ 1 claim, fr. of claims made by 1-claimers, etc.
  group_by(year, drug_group) %>%
  mutate(pct_bene=pct(n_bene/sum(n_bene)),
         pct_claims=pct(n_class_claims/sum(n_class_claims))) %>%
  select(year, drug_group, n_claims, pct_bene, pct_claims) %>%
  # group 3+ claims into one group, then sum up fractions in that new group
  mutate(n_claims=fct_other(factor(n_claims), keep=0:2, other_level='3+')) %>%
  group_by(year, drug_group, n_claims) %>%
  summarize_at(vars(pct_bene, pct_claims), sum) %>%
  ungroup() %>%
  # merge the claims group and metric into one key, and spread
  gather('key', 'val', pct_bene:pct_claims) %>%
  mutate(key2=str_c(n_claims, ':', key)) %>%
  select(year, drug_group, key2, val) %>%
  spread(key2, val) %>%
  # keep only 2011 data
  filter(year==2011) %>%
  select(-year) %>%
  kable
```

@@

ineq_summary %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  stat_smooth(aes(color=NA), color='gray', method='lm', se=F, fullrange=TRUE) +
  geom_point(shape=1) +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup}
ineq_summary %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup_by_drug}
ineq_summary %>%
  filter(unit_type=='hrr') %>%
  #rename(state=unit) %>% left_join(regions, by='state') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions) %>%
  ggplot(aes(x=fnz, y=mup, color=region)) +
  facet_wrap(~drug_group, scales='free') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='gray') +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal()
```

@@

```{r fnz_mean_region}
ineq_summary %>%
  filter(unit_type=='state', drug_group=='overall') %>%
  rename(state=unit) %>%
  filter(state != 'District of Columbia') %>%
  left_join(regions, by='state') %>%
  ggplot(aes(x=fnz, y=mean, color=region)) +
  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('All antibiotics, f> and mean consumption, states') +
  theme_minimal()
```

@@

```{r ecoli_bactrim_fnz_hrr}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_abline(intercept=16.86, slope=114.8, color='pink') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim across HRRs')
```

@@

```{r ecoli_bactrim_mup}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('mean Bactrim cons. among Bactrim-eaters') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim, mu>, HRRs')
```

@@

```{r efaecium_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. faecium') +
  ggtitle('E. faecium & beta-lactams across HRRs')

@@

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across HRRs')

@@

```{r ecoli_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams across HRRs')

@@

>>>>

```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
ineq_table = read_tsv('ineq.tsv')

ineq = ineq_table %>%
  group_by(year, unit_type, unit, drug_group) %>%
  summarize(mean=weighted.mean(n_claims, w=n_bene),
            fnz=sum(n_bene[n_claims > 0]) / sum(n_bene),
            mup=mean/fnz) %>%
  group_by(unit_type, unit, drug_group) %>%
  summarize_at(vars(mean, fnz,mup), mean) %>%
  ungroup()

@@

ineq_table %>%
  filter(unit_type=='state') %>%
  # sum across units
  group_by(year, drug_group, n_claims) %>%
  summarize(n_bene=sum(n_bene)) %>%
  ungroup() %>%
  # compute the number of claims made by people making 1 claim, etc.
  mutate(n_class_claims = n_claims * n_bene) %>%
  # compute percentages: fr. of benes w/ 1 claim, fr. of claims made by 1-claimers, etc.
  group_by(year, drug_group) %>%
  mutate(pct_bene=pct(n_bene/sum(n_bene)),
         pct_claims=pct(n_class_claims/sum(n_class_claims))) %>%
  select(year, drug_group, n_claims, pct_bene, pct_claims) %>%
  # group 3+ claims into one group, then sum up fractions in that new group
  mutate(n_claims=fct_other(factor(n_claims), keep=0:2, other_level='3+')) %>%
  group_by(year, drug_group, n_claims) %>%
  summarize_at(vars(pct_bene, pct_claims), sum) %>%
  ungroup() %>%
  # merge the claims group and metric into one key, and spread
  gather('key', 'val', pct_bene:pct_claims) %>%
  mutate(key2=str_c(n_claims, ':', key)) %>%
  select(year, drug_group, key2, val) %>%
  spread(key2, val) %>%
  # keep only 2011 data
  filter(year==2011) %>%
  select(-year) %>%
  kable
```

@@

ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  stat_smooth(aes(color=NA), color='gray', method='lm', se=F, fullrange=TRUE) +
  geom_point(shape=1) +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup}
ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r fnz_mup_by_drug}
ineq %>%
  filter(unit_type=='hrr') %>%
  #rename(state=unit) %>% left_join(regions, by='state') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions) %>%
  ggplot(aes(x=fnz, y=mup, color=region)) +
  facet_wrap(~drug_group, scales='free') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='gray') +
  geom_smooth(method='lm', se=F) +
  geom_point(shape=1) +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal()
```

@@

```{r fnz_mean_region}
ineq %>%
  filter(unit_type=='state', drug_group=='overall') %>%
  rename(state=unit) %>%
  filter(state != 'District of Columbia') %>%
  left_join(regions, by='state') %>%
  ggplot(aes(x=fnz, y=mean, color=region)) +
  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('All antibiotics, f> and mean consumption, states') +
  theme_minimal()
```

@@

```{r ecoli_bactrim_fnz_hrr}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_abline(intercept=16.86, slope=114.8, color='pink') +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim across HRRs')
```

@@

```{r ecoli_bactrim_mup}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('mean Bactrim cons. among Bactrim-eaters') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim, mu>, HRRs')
```

@@

```{r efaecium_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. faecium') +
  ggtitle('E. faecium & beta-lactams across HRRs')

@@

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across HRRs')

@@

```{r ecoli_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams across HRRs')

@@

## Using logistic models

We don't expect that, over a large range, the consumption/resistance relationship will be linear. We expect it would be something more like a logistic function. To evaluate the difference between these two approaches, I performed quasibinomial regressions $R \sim f_>$. If $\bar{f}$ is the midpoint of the $f_>$ values (i.e., the weighted mean) and $\hat{y}$ is the predicted resistance from the model, then the slope at the midpoint is $\beta_f \hat{y}(\bar{x}) \left[ 1- \hat{y}(\bar{x}) \right]$. I compare the $p$-values and slopes in the linear and quasibinomial regressions:

```{r linear_quasibin}
results %>%
  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_log'), term=='fnz') %>%
  select(bug, drug_group, model, estimate, p.value) %>%
  group_by(model) %>%
  mutate(p.value = p.adjust(p.value, method='BH')) %>%
  ungroup() %>%
  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_point() +
  geom_line(color='gray')
```

Points above the black line are significant. Note that the $p$-values in the two cases are really similar (i.e., most dumbbells are on the same horizontal line) but they can differ somewhat in their slope. The right-most pink dot is pneumococcus/macrolides, and the next is *P. mirabilis*/Bactrim.

I also tried doing beta regressions, which get similar-sized changes in the slope but really significant $p$-values. Not sure what to make of those.
